\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)

### 2015-09-09 overview.tex
modified:   casestudies.tex 	modified:   intro.tex 	modified:   methodology.tex 	modified:   overview.tex 	modified:   paper.tex 	modified:   psharp.tex
various updates

\subsection{Challenges in testing}
\label{sec:bg:challenges}

@@

A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.  A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking. Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}. A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system. Even if the model is verified, there is no guarantee that the code that will actually execute is free of bugs.

@@

Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states. Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}. Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property. Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.

>>>>

@@

@@

Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states. The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent. There was an actual bug in the system, that led this property to fail for some very rare executions. Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug until we tested vNext using the \psharp systematic testing engine.

\subsection{The \psharp framework}
\label{sec:method:psharp}

Our goal in this work is to \emph{test what is being executed}. To achieve this we use \psharp~\cite{deligiannis2015psharp}, a framework that provides: (i) an \emph{event-driven asynchronous programming} language for developing and modeling distributed systems; and (ii) a \emph{systematic concurrency testing} engine that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.

The \psharp language is an extension of \csharp, built on top of Microsoft's Roslyn\footnote{\url{https://github.com/dotnet/roslyn}} compiler, that enables asynchronous programming using communicating state-machines. \psharp machines can interact asynchronously by sending and receiving events,\footnote{We use the word ``event'' and ``message'' interchangeably.} an approach commonly used to develop distributed systems. This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g. Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).

A \psharp machine consists of an input event queue, states, state transitions, event handlers, fields and methods. Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler. This handler might update a field, create a new machine, or send an event to another machine. In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it. All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.

Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code. Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.

A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where an embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.

\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.

%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.

\subsection{Overview of our approach}
\label{sec:method:model}

In previous work~\cite{deligiannis2015psharp}, we approached the problem of testing legacy distributed systems as follows. First, we ported the system to \psharp, then we modeled its environment as \psharp state machines, and finally we tested the ported system and its environmental model using the \psharp systematic concurrency testing engine. The limitation of this approach is that it does not allow us to directly test a legacy system, as it has to be re-implemented first in \psharp. However, such endeavor is very costly and time consuming, and thus is not realistic for testing an existing production system, such as the Azure Storage vNext. Also, unless the code under test is the one that will actually execute, there is no guarantee that the real system will be bug-free.

To solve this problem, and allow \psharp to be used for testing legacy distributed systems, we decided to take a different approach. Our approach of testing existing distributed systems using \psharp requires the developer to perform three key modeling tasks:

\begin{enumerate}
\item \psharp operates in the level of communicating state machines, and thus the environment of the system-under-test must be modeled using \psharp machines, while the real components of the system must be wrapped inside \psharp machines. We call this modeled environment the \psharp test harness.

\item The top level asynchrony due to message passing, must be exposed as event sending using the \psharp APIs. If the communication layer is not based on message passing, then additional effort must be spent into refactoring the system to use message passing. This step allows the \psharp runtime to capture and control the nondeterminism due to message passing interleavings during systematic testing.

\item Any other source of nondeterminism in the system or the environment (e.g. failures and timers), must be explicitly modeled using the available \psharp APIs. This would allow \psharp to capture the nondeterminism and control it during systematic testing.
\end{enumerate}

In the remaining of this section, we will use the Azure Storage vNext case study as a running example of how to model a typical distributed system using \psharp. In principle, this modeling methodology is not specific to \psharp but can be used in combination with any systematic testing tool. We decided to use \psharp as it is a mature tool that provides a lot of modeling power via its \csharp language extensions and has an embedded systematic concurrency testing engine inside its runtime.

We argue that our approach is \emph{flexible} since it allows the user to model \emph{as much} or \emph{as little} of the environment as required to achieve the desired level of testing. We also argue that our approach is \emph{generic} since a programmer can build on top of it to test other distributed systems besides vNext (see Section~\ref{sec:exp}). Furthermore, the language features that are required to be used to connect the real code with the modeled code, are already being heavily used in production for testing purposes (e.g. \emph{virtual method dispatch}), which significantly lowers the bar for product groups to embrace \psharp for testing.

In Section~\ref{sec:psharp} we will present more specifics of how the \psharp bug-finding runtime works and extensions that we did since the original work~\cite{deligiannis2015psharp} to be able to use \psharp to systematically test production code.

### 2015-09-09 psharp.tex
various updates

Liveness property specifications are in principle harder to encode since they apply over entire program executions instead of individual program states. The \psharp developer can write liveness properties using a \emph{liveness monitor}, which is a special type of monitor that can contain three types of states: regular ones; \emph{hot} states; and \emph{cold} states. A state annotated with the \texttt{Hot} attribute denotes a state where the liveness property is not satisfied (e.g. a node has failed but a new one has not come up yet). A state annotated with the \texttt{Cold} attribute denotes a state where the liveness property is satisfied.

@@

The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent. There was an actual bug in the system, that led this property to fail for some very rare executions. Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug until we tested vNext using the \psharp systematic testing engine.

>>>>

Liveness property specifications are in principle harder to encode since they apply over entire program executions instead of individual program states. Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}. Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property. Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.

The \psharp developer can write liveness properties using a \emph{liveness monitor}, which is a special type of monitor that can contain three types of states: regular ones; \emph{hot} states; and \emph{cold} states. A state annotated with the \texttt{Hot} attribute denotes a state where the liveness property is not satisfied (e.g. a node has failed but a new one has not come up yet). A state annotated with the \texttt{Cold} attribute denotes a state where the liveness property is satisfied.

@@

### 2015-09-09 related.tex
various updates

>>>>

A completely different approach for reasoning about the correctness of distributed systems is to use formal methods.  A notable example is TLA+~\cite{lamport1994temporal}, a formal specification language that can be used to design and verify concurrent programs via model checking. Amazon recently published an article describing their use of TLA+ in Amazon Web Services to verify distributed protocols~\cite{newcombe2015aws}. A limitation of TLA+, as well as other similar specification languages, is that they are applied on a model of the system and not the actual system. Even if the model is verified, the gap between a real-world implementation and the verified model is still significant, so implementation bugs are still a realistic concern.
